Staged typed type-preserving tagless ``interpreters'', \emph{finally}

### 2007-04-04 tagless/tagless-final.tex
Move figures closer to exposition

This solution using GADTs works but is not quite satisfactory. First, it
cannot be ported to MetaOCaml as GADTs are unavailable there.  Second,
the problem of nonexhaustive pattern\hyp matching reappears in |app|
above: the type |P t| has four constructors, of which pattern\hyp
matching in |app| uses only |VF| and~|E|. One may say that the
constructors |VI| and |VB| obviously cannot occur because they do not
construct values of type |P (a -> b)| as required by the type of |app|.
Indeed the compiler could reason thus, but it may not; GHC for one
issues warnings.  Although this point may seem minor, it is the heart of
the tagging problem and the purpose of tag elimination. A typed tagged
interpreter contains many pattern\hyp matching forms that look partial
but are total in reality. The goal is to make this totality
\emph{syntactically} apparent.

@@

@@

Unlike the GADT approach in~\S\ref{S:PE-GADT}, all pattern\hyp matching
in~|P| is \emph{syntactically} exhaustive, so it is \emph{patent} to the
compiler that |P| is total.  Further, all pattern\hyp matching occurs
during partial evaluation, only to check if a value is known statically,
never what type it has.  In other words, our partial evaluator tags
phases (with |Some| and |None|) but not object types.

>>>>

@@

This solution using GADTs works but is not quite satisfactory. First, it
cannot be ported to MetaOCaml as GADTs are unavailable there.  Second,
the problem of nonexhaustive pattern\hyp matching reappears in |app|
above: the type |P t| has four constructors, of which pattern\hyp
matching in |app| uses only |VF| and~|E|. One may say that the
constructors |VI| and |VB| obviously cannot occur because they do not
construct values of type |P (a -> b)| as required by the type of |app|.
Indeed the metalanguage implementation could reason thus, but it may not; GHC for one
issues warnings.  Although this point may seem minor, it is the heart of
the tagging problem and the purpose of tag elimination. A typed tagged
interpreter contains many pattern\hyp matching forms that look partial
but are total in reality. The goal is to make this totality
\emph{syntactically} apparent.

@@

Unlike the GADT approach in~\S\ref{S:PE-GADT}, all pattern\hyp matching
in~|P| is \emph{syntactically} exhaustive, so it is patent to the metalanguage
implementation that |P| is total.  Further, all pattern\hyp matching occurs
during partial evaluation, only to check if a value is known statically,
never what type it has.  In other words, our partial evaluator tags
phases (with |Some| and |None|) but not object types.
